---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Lindsey Wilson"
date: "2/7/23"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
dn <- dennys
lq <- laquinta
```

### Exercise 1

The code below gives the number of Denny's locations in Alaska:

```{r dennys-in-alaska}
dn_ak <- dn %>%
  filter(state == "AK")
message("There are ", nrow(dn_ak), " Denny's locations in Alaska")
```

### Exercise 2

And this tells us how many La Quinta locations are in Alaska: 

```{r laquinta-in-alaska}
lq_ak <- lq %>%
  filter(state == "AK")
message("There are ", nrow(lq_ak), " La Quinta locations in Alaska")
```

### Exercise 3

Since there are 3 Denny's and 2 La Quinta's in Alaska, if we want to calculate the distances between each pair of Denny's and La Quinta then in total we'll need to calculate 6 distances.

### Exercise 4

Below we've used `full_join()` to put together the `dn_ak` and `lq_ak` datasets:

```{r full-join}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak
```
This code tells us how many observations are in our newly created dataset, along with the names of the columns

```{r}
message("There are ", nrow(dn_lq_ak), " observations in the dn_lq_ak dataset")
names(dn_lq_ak)
```

### Exercise 5

It looks like each row represents a Denny's/La Quinta pair, so to get the distances between each pair we can just calculate a distance column for each row. We would do this with the `mutate()` function.

### Exercise 6

This code creates a function to calculate distance for each row:

```{r distance function}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```

Which allows us to calculate a distance variable for each row in `dn_lq_ak`:

```{r calculate-distance}
dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x,
                              latitude.x,
                              longitude.y,
                              latitude.y
                              ))
```

### Exercise 7

The code below generates a data frame that has one observation for each Denny's location, along with the distance to the closest La Quinta:

```{r minimum-distance}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
```

### Exercise 8

We can use the data frame we generated in exercise 7 to summarize the distribution of distances between Denny's and the nearest La Quinta:

```{r min-distance-summary}
ggplot(data = dn_lq_ak_mindist, mapping = aes(x = closest)) + 
  geom_histogram(aes(y = after_stat(density)),
                     binwidth = 2, 
                     fill = "Blue") + 
  geom_density()

message("Mean = ", mean(dn_lq_ak_mindist$closest), " miles.")
message("SD = ", sd(dn_lq_ak_mindist$closest), " miles.")
```

With only 3 Denny's locations we can't say much about the shape of the distribution, but the descriptives at least make it seem like Denny's and La Quinta locations are pretty close together.

### Exercise 9

We're now going to repeat this analysis pipeline for the state of North Carolina.
The code to do that is below:

```{r NC-analysis}
#Filters Denny's for NC locations
dn_nc <- dn %>%
  filter(state == "NC")

#Filter La Quinta for NC locations
lq_nc <- lq %>%
  filter(state == "NC")

#Join dn_nc and lq_nc datasets by state
dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")

#Calculate distances between all Dennys-LaQuinta pairs
dn_lq_nc <- dn_lq_nc %>%
  mutate(distance = haversine(longitude.x,
                              latitude.x,
                              longitude.y,
                              latitude.y
                              ))

#Find minimum distance to a La Quinta for each Denny's
dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

#Histogram and descriptives of minimum distances 
#(horrible yellow and purple graph is on purpose)
ggplot(data = dn_lq_nc_mindist, mapping = aes(x = closest)) + 
  geom_histogram(aes(y = after_stat(density)),
                 binwidth = 30, 
                 fill = "Yellow") + 
  geom_density() +
  theme(panel.background = element_rect(fill = "Purple"))

message("Mean = ", mean(dn_lq_nc_mindist$closest), " miles.")
message("SD = ", sd(dn_lq_nc_mindist$closest), " miles.")

```

Right away we can see from this analysis that there are many more Denny's locations in NC than in Alaska, and that on average they're much further away from the nearest La Quinta.
However, it doesn't really feel valid to compare the two states because, again, there were only 3 Denny's locations in Alaska at all.

### Exercise 10

Let's repeat the same analysis for Texas:

```{r TX-analysis}
#Filters Denny's for TX locations
dn_tx <- dn %>%
  filter(state == "TX")

#Filter La Quinta for TX locations
lq_tx <- lq %>%
  filter(state == "TX")

#Join dn_tx and lq_tx datasets by state
dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")

#Calculate distances between all Dennys-LaQuinta pairs
dn_lq_tx <- dn_lq_tx %>%
  mutate(distance = haversine(longitude.x,
                              latitude.x,
                              longitude.y,
                              latitude.y
                              ))

#Find minimum distance to a La Quinta for each Denny's
dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

#Histogram and descriptives of minimum distances 
ggplot(data = dn_lq_tx_mindist, mapping = aes(x = closest)) + 
  geom_histogram(aes(y = after_stat(density)),
                 binwidth = 3, 
                 fill = "Blue") + 
  geom_density()
  
message("Mean = ", mean(dn_lq_tx_mindist$closest), " miles.")
message("SD = ", sd(dn_lq_tx_mindist$closest), " miles.")

```

### Exercise 11

And for California:

```{r CA-analysis}
#Filters Denny's for CA locations
dn_ca <- dn %>%
  filter(state == "CA")

#Filter La Quinta for CA locations
lq_ca <- lq %>%
  filter(state == "CA")

#Join dn_ca and lq_ca datasets by state
dn_lq_ca <- full_join(dn_ca, lq_ca, by = "state")

#Calculate distances between all Dennys-LaQuinta pairs
dn_lq_ca <- dn_lq_ca %>%
  mutate(distance = haversine(longitude.x,
                              latitude.x,
                              longitude.y,
                              latitude.y
                              ))

#Find minimum distance to a La Quinta for each Denny's
dn_lq_ca_mindist <- dn_lq_ca %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

#Histogram and descriptives of minimum distances 
ggplot(data = dn_lq_ca_mindist, mapping = aes(x = closest)) + 
  geom_histogram(aes(y = after_stat(density)),
                 binwidth = 5, 
                 fill = "Blue") + 
  geom_density()
  
message("Mean = ", mean(dn_lq_ca_mindist$closest), " miles.")
message("SD = ", sd(dn_lq_ca_mindist$closest), " miles.")

```

### Exercise 12

From these analyses, it looks like Texas is the state where the Mitch Hedberg joke seems the most applicable.
It has enough observations to be able to draw solid conclusions about it (unlike Alaska), and it has both the smallest mean Dennys-to-La Quinta distance and the smallest standard deviation.
The distribution of minimum distances is also extremely positively skewed (not unlike the others), meaning that the actual center of the minimum distance distribution is likely somewhat smaller than the mean is indicating.

Just to drive that intuition home, we can calculate the median distances below:

```{r medians}
message("NC median = ", median(dn_lq_nc_mindist$closest), " miles") +
message("TX median = ", median(dn_lq_tx_mindist$closest), " miles") +
message("CA median = ", median(dn_lq_ca_mindist$closest),  " miles")
```

So. there we have it. 
"La Quinta" is Spanish for "next to a Dennys", but that's especially true in Texas. 