---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Lindsey Wilson"
date: "2/7/23"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
dn <- dennys
lq <- laquinta
```

### Exercise 1

The code below gives the number of Denny's locations in Alaska:

```{r dennys-in-alaska}
dn_ak <- dn %>%
  filter(state == "AK")
message("There are ", nrow(dn_ak), " Denny's locations in Alaska")
```

### Exercise 2

And this tells us how many La Quinta locations are in Alaska: 

```{r laquinta-in-alaska}
lq_ak <- lq %>%
  filter(state == "AK")
message("There are ", nrow(lq_ak), " La Quinta locations in Alaska")
```

### Exercise 3

Since there are 3 Denny's and 2 La Quinta's in Alaska, if we want to calculate the distances between each pair of Denny's and La Quinta then in total we'll need to calculate 6 distances.

### Exercise 4

Below we've used `full_join()` to put together the `dn_ak` and `lq_ak` datasets:

```{r full-join}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak
```
This code tells us how many observations are in our newly created dataset, along with the names of the columns

```{r}
message("There are ", nrow(dn_lq_ak), " observations in the dn_lq_ak dataset")
names(dn_lq_ak)
```

### Exercise 5

It looks like each row represents a Denny's/La Quinta pair, so to get the distances between each pair we can just calculate a distance column for each row. We would do this with the `mutate()` function.

### Exercise 6

This code creates a function to calculate distance for each row:

```{r distance function}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```

Which allows us to calculate a distance variable for each row in `dn_lq_ak`:

```{r calculate-distance}
dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x,
                              latitude.x,
                              longitude.y,
                              latitude.y
                              ))
```










